---
title: "Text mining emotion statements"
output:
  html_document:
    df_print: paged
---
This script uses the package sentimentr to assess the emotional content
in sentences about the prospects related to a rival candidate would win
the presidential election

It returns a dataframe where each row represents an emotion detected
within their response. It displays the total number of words on their
submission, and the number of emotion phrases detected in it.

Additionally, it provides an emotional rate per sentence, which is then
averaged in the whole submission. This emotional rate is based in a basic
dictionary lookup to find the emotion words. This rate ranges from 0 to 1.
The dictionary used is a filtered version of the NRC emotion lexicon from
Mohammad & Turney (2010).

```{r setup}
require(tidyverse)
require(ggplot2)
require(sentimentr)

folder_root <- dirname(dirname(rstudioapi::getActiveDocumentContext()$path))
folder_data <- file.path(folder_root, "Data")

```

Here we load the base data
```{r}
# Load base data
raw_data <- read.csv(file = file.path(folder_data, "emotion_prime_scoring.csv")) %>% 
  select(1:3) %>% 
  rename("subject" = 1, "candidate" = 2, "response" = 3) %>% 
  na.omit() %>% 
  filter(response != "")

```

Now we divide the statements in sentences, and perform the emotion detection for each subject
```{r}
emotional_data <- raw_data %>%
  get_sentences() %>% 
  emotion_by(response, by = c("subject", "candidate")) %>% 
  filter(emotion_count != 0) %>% 
  mutate(candidate = as.factor(candidate),
         emotion_type = as.factor(emotion_type))

emotional_data

```

Now we create summaries per candidate
```{r}
candidate_data <- emotional_data %>% 
  group_by(candidate) %>% 
  mutate(members = length(unique(subject))) %>% 
  ungroup() %>% 
  group_by(candidate, emotion_type, members) %>% 
  summarise(n_emotion = n(), proportion = n_emotion/members) %>% 
  ungroup() %>% 
  unique() %>% 
  mutate(emotion_type = fct_reorder(emotion_type, proportion)) %>% 
  filter(!str_detect(emotion_type, "negated"))

candidate_data
  

```

Quick visualization
```{r fig.width=10}
ggplot(data = candidate_data) +
  geom_col(aes(x = emotion_type,
               y = proportion,
               fill = emotion_type)) +
  
  facet_grid(.~candidate) +
  
  theme_bw() +
  
  scale_y_continuous(name = "Proportion of participants expressing emotion") +
  
  scale_fill_brewer(palette = "Dark2")
```

























